---
title: "GG Plot 1"
author: "Kylie Wheelock Riley"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```

## Create weather data from NOAA 
```{r}
weather_df = 
  ## Dowloading wather from 3 stations
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Create a gg plot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
##need to add  + geom(point) to get the data to plot

##alternate way of making this plot using piping (if you need to modify data before it)
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

## Saving 
```{r}
## below is code to save the output and then plot it later.  Not generally the best workflow
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))


```

## Advanced Scatterplot
```{r}
##Using name as the color aestetic 
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))

##Adding a smooth curve with geom_smooth
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

## Facets
```{r}
##separating out variables by faceting them into their own plot
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

```{r}
## looking at tmax by date, and points are precipitation
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

```

## Learning Assessment 1 
Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax_far = (tmax * (9/5) + 32),
    tmin_far = (tmin * (9/5) + 32)
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) 
```



```{r}
## Line plot showing max temp by date for each location 
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 

## More things - The geoms that you choose are independent and can be what you think are most useful. The graph above only uses geom smooth, not geom point to display the data. 

```

## 2D density
Hex is good for not showing all of the data in large datasets, avoids overlapping
```{r}
##can usegeom_hex if you install (install.packages("hexbin))
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_bin2d()
```

## Learning Assessment 2
In the preceding, we set the alpha aesthetic “by hand” instead of mapping it to a variable. This is possible for other aesthetics too. To check your understanding of this point, try to explain why the two lines below don’t produce the same result:

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))

## In the second plot, color is a variables and ggplot is giving the plot a default color.  In the first you are assigning the plot the color blue. 
```

## Univariate plots
```{r}
##Histogram
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()

##Playing around with bin width and colors - dodge separates the bars so they aren't stacked 
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
  
```


```{r}
##Density plots - transparensy set to 0.4 so all values appear
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") 
  
```

## Box plots
```{r}
weather_df %>% 
  ggplot(aes( x = name, y = tmax)) +
  geom_boxplot()
```

## Violin plots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)
```

## Ridge plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(scale = .85)
  
```

## Learning assessment 3
Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.
```{r}
##Histogram
ggplot(weather_df, aes(x = prcp, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)

##Density plot
ggplot(weather_df, aes(x = prcp, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") 

##Box plot
weather_df %>% 
  ggplot(aes( x = name, y = prcp)) +
  geom_boxplot()

##Violin plot
ggplot(weather_df, aes(x = name, y = prcp)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)

#Ridge plot
weather_df %>% 
  ggplot(aes(x = prcp, y = name)) +
  geom_density_ridges(scale = 1)
## want to zoom in on this one to see differences, not sure how to do that

##data is highly skewed, winner is boxplot either omitting days with zero precip, or capping precip at 100:
weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)


```


# Saving and embedding plots

## GG Save
```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
  
```

## Embedding a plot
```{r, fig.width = 6}

## Good default settings
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) 

```

```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) 

```

## Learning assessment 4
Set global options for your figure sizes in the “setup” code chunk and re-knit your document. What happens when you change fig.asp? What about out.width?

