---
title: "GG Plot 1"
author: "Kylie Wheelock Riley"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```

## Create weather data 
```{r}
weather_df = 
  ## Dowloading wather from 3 stations
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Create a gg plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

##alternate way of making this plot
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

## Saving initial plots
```{r}
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

## Adding color

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4)
```

## Why do aestetic "aes" positions matter?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(se = FALSE)
```

## Facets

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(se = FALSE) +
  facet_grid(~name)
```

## More meaningful graphs
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.35) +
  geom_smooth(size = 2, se = FALSE)
```

## Do Learning Assessment 1 here 

## More things
THe geoms that you choose are independent and can be what you think are most useful 
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(size = 2, se = FALSE)
```


## 2D density
Hex is good for not showing all of the data in large datasets
```{r}
##can usegeom_hex if you install (install.packages("hexbin))
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_bin2d()
```

## More plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(~name)
```

density plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
   geom_density(alpha = .3) 
  
```

boxplot
```{r}
weather_df %>% 
  ggplot(aes( x = name, y = tmax)) +
  geom_boxplot()
```

violin plot
```{r}
weather_df %>% 
  ggplot(aes( x = name, y = tmax)) +
  geom_violin()
```

ridge plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(scale = .85)
  
```

## Saving a plot

```{r}
ggp_ridge_temp =
  weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_ridge_temp)
  
```

imbedding a plot
```{r, fig.width = 6}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) 

```

```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) 
```


