---
title: "GG Plot 2"
author: "Kylie Wheelock Riley"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
```


## Create weather data 
```{r load_data, cache = TRUE}
##Cache only runs this dataset once so each time you knit it is not pulling it from the server. 

weather_df = 
  ## Dowloading wather from 3 stations
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Adding labels to a gg plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )

```

## Location of ticks on axes using scale_x_ and scale_y_
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15"))

##the version below also scales the y axis using a sqrt, good for log tansformed vars
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

## Colors and themes using scale_color_hue
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) + 
  labs(
    title = "Temperature Plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
    ) +
  scale_color_hue(
    name = "Weather Station",
    h = c(100, 200)
  )
      
```

## Using viridis for more specific/ specialized color schemes
```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_temp_plot

## The sample above uses the defauly viridis scheme, can choose more specific ones
## used discrete = TRUE because the color aesthetic is mapped to a discrete variable. For continuous variables you can omit this. 
```

## Themes 

Changes general organization

```{r}
##generally good to plot legend on the bottom, so you have more room for your actual graph
ggp_temp_plot + 
  theme(legend.position = "bottom")

```

### bw theme, not black and white
```{r}
ggp_temp_plot + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

### Classic theme
```{r}
ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

### Minimal plot theme
```{r}
ggp_temp_plot + 
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Learning Assessment 1
Revisit the plot showing tmax against date for each location. Use labels, scale options, and theme changes to improve the readability of this plot.

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Maximum Temperature plot",
    x = "Date",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) +
    theme_minimal() +
  theme(legend.position = "bottom")

```

# Setting Options

```{r}
##Use this code at the topof every .rmd file to keep figure formatting consistent
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Data Arguments in geom
Better understandingofhow each layer / piece of code works in an example that splits weather_df into separate datasets for Central Park and Waikiki 
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)

```


brief aside about colors
```{r}
waikiki %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(alpha = .5, color = "blue")
```


## Patchwork - Multi-panel plots 

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

##(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

# Data Manipulation
tidy data allows you to make nicer grapha
setting it as a factor allows you to have nicer organization 

## Ordering factor Variables 
```{r}
##You can change the order level of a factor variable to your specified preference using forcats::fct_relevel or according to the value of another variable using forcats::fct_reorder.

## Reordering by hand
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")

## Reordering name according to tmax values in each name
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

## Separating into separate plots (tmax and tmin by location)
```{r}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```

## Using pitvot_longer to rearrange data for plot
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(visit = recode(visit, "bl" = "00m"),
         visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```



