---
title: "GG Plot 2"
author: "Kylie Wheelock Riley"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
```


## Create weather data 
```{r load_data, cache = TRUE}
##Cache only runs this dataset once so each time you knit it is not pulling it from the server. 

weather_df = 
  ## Dowloading wather from 3 stations
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Adding labels to a gg plot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) + 
  labs(
    title = "Temperature Plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5C", "20C")
  ) +
  ##taking sqrt, good for non nomally distributed data
  ##Not very useful here because it drops all of the negative data
  scale_y_continuous(
    trans = "sqrt"
  )


```

## Choosing colors in gg plot using scale_color_hue
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) + 
  labs(
    title = "Temperature Plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
    ) +
  scale_color_hue(
    name = "Weather Station",
    h = c(100, 200)
  )
      
```

## viridis package good for color blind
```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
    ##variable used is a discrete variable, do ned to specify that it is true
  )
```

## Themes 

Changes general organizarion

```{r}
##generally good to plot legend on the bottom, so you have more room for your actual graph
ggp_temp_plot + 
  theme(legend.position = "bottom")

```
### Black and white plot
```{r}
ggp_temp_plot + 
  theme_bw() +
  theme(legend.position = "bottom")
```
### Minimal plot theme
```{r}
ggp_temp_plot + 
  theme_minimal()+
  theme(legend.position = "bottom")
```

where do warning come from
```{r}
weather_df %>% 
  filter(is.na(tmin))
```

## More than 1 dataset
Figure out why not working after class
```{r}
central_park = 
  weather_df %>% 
  filter(name == "Central_Park_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HI")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp)) +
  geom_line(data = central_park)

```

brief aside about colors
```{r}
waikiki %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(alpha = .5, color = "blue")
```


## Multi-panel plots 
patchwork package, install after class

```{r}
ggp_scatter =
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_density = 
  weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_density()


```

